Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-15 paper/sections/preliminaries.tex
[ paper ] citation using natbib.
[ travis-ci ] fixed cpphs requirement.

\textit{Propositions.}
We define the classical propositional logic as usual.
A proposition is an expression of indivisible propositional atoms
(e.g., symbols $φ₀, φ₁, \cdots$),
binary logic connectives ($∧, ∨, ⇒, ⇔)$,
propositional negation $(¬)$, and logic constants $⊥$
and $⊤$. We use the inductive definition of the set \PROP
presented in \cite{VanDalen1994} as the mechanism to produce any
propositional formula.
Then, $ϕ_i \in \PROP$ for all $i \in \mathbb{N}$.
If $φ, ψ\in \PROP$, then $(φ\ \square\ ψ) \in \PROP$,
where $(φ\ \square\ ψ)$ stands for a proposition that use a
binary connective in the place for the $\square$ symbol.
If $φ \in \PROP$, then $(¬ φ) \in \PROP$.
Finally, $⊥, ⊤ \in \PROP$.

@@

\emph{Natural Deductions.}
The natural deduction designed by Gentzen \improvement{Cite the work
by Gentzen.} defines a formal system for derivations that uses a set
of inference rules as the only way to obtain a conclusion.
We use this system to refer to formal proofs as natural deduction
proofs and conversely, in the latter.
We define our inference rules for \CPL in the Fig.
\ref{fig:CPL-inference-rules}. These rules give us the complete
meaning of the connectives; how to introduce them and
how they can be remove them in the conclusion side of the theorems.

\begin{figure}
\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{}
\RightLabel{\scriptsize\tt assume $φ$}
\UnaryInfC{$Γ , φ ⊢ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\scriptsize\tt weaken $ψ$}
\UnaryInfC{$Γ , ψ ⊢ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{}
\RightLabel{\scriptsize\tt $⊤$-intro}
\UnaryInfC{$Γ ⊢ ⊤$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ ⊥$}
\RightLabel{\scriptsize\tt $⊥$-elim $φ$}
\UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
%}
\]

\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ , φ ⊢ ⊥$}
\RightLabel{\scriptsize\tt $¬$-intro}
\UnaryInfC{$Γ ⊢ ¬ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ ¬ φ$}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\scriptsize\tt $¬$-elim}
\BinaryInfC{$Γ ⊢ ⊥$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\AxiomC{$Γ ⊢ ψ$}
\RightLabel{\scriptsize\tt $\wedge$-intro}
\BinaryInfC{$Γ ⊢ φ \wedge ψ$}
\end{bprooftree}
%}
\]

\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \wedge ψ$}
\RightLabel{\scriptsize\tt $\wedge$-proj$_1$}
\UnaryInfC{$Γ⊢ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \wedge ψ$}
\RightLabel{\scriptsize\tt $\wedge$-proj$_2$}
\UnaryInfC{$Γ⊢ ψ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\scriptsize\tt $\vee$-intro$_1$ $ψ$}
\UnaryInfC{$Γ ⊢ φ \vee ψ$}
\end{bprooftree}
%}
\]

\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ ψ$}
\RightLabel{\scriptsize\tt $\vee$-intro$_2$ $φ$}
\UnaryInfC{$Γ ⊢ φ \vee ψ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ,φ ⊢ \gamma $}
\AxiomC{$Γ , ψ  ⊢ \gamma$}
\RightLabel{\scriptsize\tt $\vee$-elim}
\BinaryInfC{$Γ , φ \vee ψ ⊢ \gamma$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ , φ ⊢ ψ$}
\RightLabel{\scriptsize\tt $\Rightarrow$-intro}
\UnaryInfC{$Γ ⊢ φ \Rightarrow ψ$}
\end{bprooftree}
%}
\]

\[
%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \Rightarrow ψ$}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\scriptsize\tt $\Rightarrow$-elim}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ , φ ⊢ ψ$}
\AxiomC{$Γ , ψ ⊢ φ$}
\RightLabel{\scriptsize\tt $⇔$-intro}
\BinaryInfC{$Γ ⊢ φ ⇔ ψ$}
\end{bprooftree}
%}
\]

\[
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\AxiomC{$Γ ⊢ φ ⇔ ψ$}
\RightLabel{\scriptsize\tt $⇔$-elim$_1$}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}
\qquad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\AxiomC{$Γ ⊢ φ ⇔ ψ$}
\RightLabel{\scriptsize\tt $⇔$-elim$_2$}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}
\]

>>>>

\textit{Propositions.}
We define the classical propositional logic as usual.
A proposition is an expression of indivisible propositional atoms
(e.g., symbols $φ₀, φ₁, \cdots$),
binary logic connectives ($∧, ∨, ⇒, ⇔)$,
propositional negation $(¬)$, and logic constants $⊥$
and $⊤$. We use the inductive definition of the set \Prop
presented in \cite{VanDalen1994} as the mechanism to produce any
propositional formula.
Then, $ϕ_i \in \Prop$ for all $i \in \mathbb{N}$.
If $φ, ψ\in \Prop$, then $(φ\ \square\ ψ) \in \Prop$,
where $(φ\ \square\ ψ)$ stands for a proposition that use a
binary connective in the place for the $\square$ symbol.
If $φ \in \Prop$, then $(¬ φ) \in \Prop$.
Finally, $⊥, ⊤ \in \Prop$.

@@

\emph{Natural Deductions.}
The natural deduction designed by Gentzen \improvement{Cite the work
by Gentzen.} defines a formal system for derivations that uses a set
of inference rules as the only way to obtain a conclusion.
We use this system to refer to formal proofs as natural deduction
proofs and conversely, in the latter.
We define our inference rules for \CPL in the Fig.
\ref{fig:CPL-inference-rules}. These rules give us the complete
meaning of the connectives; how to introduce them and
how they can be remove them in the conclusion side of the theorems.
\change{the size of the text.}

\begin{figure}
\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{}
\RightLabel{\tt assume $φ$}
\UnaryInfC{$Γ , φ ⊢ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\tt weaken $ψ$}
\UnaryInfC{$Γ , ψ ⊢ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{}
\RightLabel{\tt $⊤$-intro}
\UnaryInfC{$Γ ⊢ ⊤$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ ⊥$}
\RightLabel{\tt $⊥$-elim $φ$}
\UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
%}
\]

\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ , φ ⊢ ⊥$}
\RightLabel{\tt $¬$-intro}
\UnaryInfC{$Γ ⊢ ¬ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ ¬ φ$}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\tt $¬$-elim}
\BinaryInfC{$Γ ⊢ ⊥$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\AxiomC{$Γ ⊢ ψ$}
\RightLabel{\tt $\wedge$-intro}
\BinaryInfC{$Γ ⊢ φ \wedge ψ$}
\end{bprooftree}
%}
\]

\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \wedge ψ$}
\RightLabel{\tt $\wedge$-proj$_1$}
\UnaryInfC{$Γ⊢ φ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \wedge ψ$}
\RightLabel{\tt $\wedge$-proj$_2$}
\UnaryInfC{$Γ⊢ ψ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\tt $\vee$-intro$_1$ $ψ$}
\UnaryInfC{$Γ ⊢ φ \vee ψ$}
\end{bprooftree}
%}
\]

\[%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ ψ$}
\RightLabel{\tt $\vee$-intro$_2$ $φ$}
\UnaryInfC{$Γ ⊢ φ \vee ψ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ ,φ ⊢ \gamma $}
\AxiomC{$Γ , ψ  ⊢ \gamma$}
\RightLabel{\tt $\vee$-elim}
\BinaryInfC{$Γ , φ \vee ψ ⊢ \gamma$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ , φ ⊢ ψ$}
\RightLabel{\tt $\Rightarrow$-intro}
\UnaryInfC{$Γ ⊢ φ \Rightarrow ψ$}
\end{bprooftree}
%}
\]

\[
%\scalebox{0.9}{
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ \Rightarrow ψ$}
\AxiomC{$Γ ⊢ φ$}
\RightLabel{\tt $\Rightarrow$-elim}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}\quad
\begin{bprooftree}
\AxiomC{$Γ , φ ⊢ ψ$}
\AxiomC{$Γ , ψ ⊢ φ$}
\RightLabel{\tt $⇔$-intro}
\BinaryInfC{$Γ ⊢ φ ⇔ ψ$}
\end{bprooftree}
%}
\]

\[
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\AxiomC{$Γ ⊢ φ ⇔ ψ$}
\RightLabel{\tt $⇔$-elim$_1$}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}
\qquad
\begin{bprooftree}
\AxiomC{$Γ ⊢ φ$}
\AxiomC{$Γ ⊢ φ ⇔ ψ$}
\RightLabel{\tt $⇔$-elim$_2$}
\BinaryInfC{$Γ ⊢ ψ$}
\end{bprooftree}
\]

### 2017-09-15 paper/sections/proof-reconstruction.tex
[ travis-ci ] fixed cpphs requirement.

To improve that definition for terminating, we will use the
technique proposed by Bove and Capettra in
\cite{Bove2005}, which in particular with functions that
transform propositions consists of:
take a general recursive function $f: \PROP \to \PROP$.
and a predicate $P: \PROP \to \abbre{Ind}$, where $\abbre{Ind}$
is a type with a structured inductive definition.

Then, we redefine the function as
$f: \PROP \to \abbre{Ind}\ \to \PROP$, and perform the recursion
using induction on the structure of the \abbre{Ind} type.

@@

Thus, we have a new function
$\unshunt_{1} : \PROP \to \Nat \to \PROP$
defined by structural recursion in its second argument in
Eq.~\ref{eq:unshunt-structured}.

@@

A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as well as \emph{the rank for a
proposition} is a function with type
$\PROP \to \Nat$ that assigns a natural number to a
proposition φ as follows.

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
rank(φ₁) + c_{¬}  & φ ≡ \neg φ₁, \\
\max{(rank(φ₁),rank(φ₂)} + c_{\square}
  & φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

@@

Therefore, we define an appropriate complexity measure
$\unshunt_{cm}: \PROP → \NN$
for the second argument of the $\unshunt_{1}$ function in
Eq.~\ref{eq:unshunt-rank-definition}.

@@

\newcommand{\thmunshunt}{\tt thm-$\unshunt$}
\begin{theorem}[\thmunshunt]
\label{thm:unshunt}
$Γ ⊢ \unshunt(φ)$, then $Γ ⊢ φ$.
\end{theorem}

@@

\newcommand{\thmsplit}{\name{thm-split}}
\begin{theorem}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\end{theorem}

@@

\emph{Resolve.}
\newcommand{\thmresolve}{\name{thm-resolve}}
The \resolve rule is a generalization version for resolution
in propositional logic since the formulas are not shaped always as
the theorem expect (see Fig.~\ref{fig:metis-inferences}).
Since our approach is mere syntactic, in order to apply such a
theorem, we perform rearrangements inside the formulas to match with
the pattern expected by the resolution theorem. If this strategy
fails at some point, we could not check the rest of the proof.

@@

\begin{itemize}
  \item \name{reorder-∨.} Given two formulas, φ and ψ, both disjunctions, fix the order of the first one by following the syntactically
  order of the second one. If the process succeed (denote by φ ⤳ ψ),
  returns the target formula, ψ, otherwise, keep the first one, φ.
  \begin{align}
  \begin{split}
&\name{reorder-∨}\ :\ \PROP \to \PROP \to \PROP\\
&\name{reorder-∨}(φ, ψ) =
  \begin{cases}
  ψ   & φ ⤳ ψ\\
  ϕ   & \text{otherwise.}
  \end{cases}
  \end{split}
  \end{align}

>>>>

To improve that definition for terminating, we will use the
technique proposed by Bove and Capettra in
\cite{Bove2005}, which in particular with functions that
transform propositions consists of:
take a general recursive function $f: \Prop \to \Prop$.
and a predicate $P: \Prop \to \abbre{Ind}$, where $\abbre{Ind}$
is a type with a structured inductive definition.

Then, we redefine the function as
$f: \Prop \to \abbre{Ind}\ \to \Prop$, and perform the recursion
using induction on the structure of the \abbre{Ind} type.

@@

Thus, we have a new function
$\unshunt_{1} : \Prop \to \Nat \to \Prop$
defined by structural recursion in its second argument in
Eq.~\ref{eq:unshunt-structured}.

@@

A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as well as \emph{the rank for a
proposition} is a function with type
$\Prop \to \Nat$ that assigns a natural number to a
proposition φ as follows.

\begin{equation*}
\label{eq:rank-definition}
rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
rank(φ₁) + c_{¬}  & φ ≡ \neg φ₁, \\
\max{(rank(φ₁),rank(φ₂))} + c_{\square}
  & φ ≡ φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}

@@

Therefore, we define an appropriate complexity measure
$\unshunt_{cm}: \Prop → \NN$
for the second argument of the $\unshunt_{1}$ function in
Eq.~\ref{eq:unshunt-rank-definition}.

@@

\begin{theorem}[\thmunshunt]
\label{thm:unshunt}
$Γ ⊢ \unshunt(φ)$, then $Γ ⊢ φ$.
\end{theorem}

@@

\begin{theorem}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\end{theorem}

@@

\emph{Resolve.}

The \resolve rule is a generalization version for resolution
in propositional logic since the formulas are not shaped always as
the theorem expect (see Fig.~\ref{fig:metis-inferences}).
Since our approach is mere syntactic, in order to apply such a
theorem, we perform rearrangements inside the formulas to match with
the pattern expected by the resolution theorem. If this strategy
fails at some point, we could not check the rest of the proof.

@@

\begin{itemize}
  \item \name{reorder-∨.} Given two formulas, φ and ψ, both disjunctions, fix the order of the first one by following the syntactically
  order of the second one. If the process succeed (denote by φ ⤳ ψ),
  returns the target formula, ψ, otherwise, keep the first one, φ.
  \begin{align}
  \begin{split}
&\name{reorder-∨}\ :\ \Prop \to \Prop \to \Prop\\
&\name{reorder-∨}(φ, ψ) =
  \begin{cases}
  ψ   & φ ⤳ ψ\\
  ϕ   & \text{otherwise.}
  \end{cases}
  \end{split}
  \end{align}
